---
title: "Assignment: personal activity monitoring"
author: "Jeroen van der A"
date: "May 14, 2016"
output: html_document
---

## calling necessary libraries
```{R libraries}
library(plyr)
library(ggplot2)
library(lubridate)



```

## loading and cleaning data
```{r loadingData}
stepdata = read.csv('activity.csv', na.strings=c("NA"))
```

## What is mean total number of steps taken per day?

```{r showTotals}
aggrsteps <- aggregate(steps ~ date, data=stepdata, sum)
hist(aggrsteps$steps, 
     xlab='steps', 
     main='Frequency of days where a number of steps have been made',
     breaks=8,
     col="lightgray"
     )

```

## Calculate and report the mean and median total number of steps taken per day

```{r calculateMeanMedian}
mean(aggrsteps$steps)
median(aggrsteps$steps)
```
#### Extra step to visualize the last question and see how the data is spread.
```{R boxplot}

boxplot(aggrsteps$steps,data=aggrsteps, 
        main="number of steps per day")
points(mean(aggrsteps$steps),
       col="red",
       pch=18)

```

## What is the average daily activity pattern?

```{R meanInterval}
intervalmean = aggregate(steps ~ interval, data=stepdata, mean)
plot(intervalmean$interval, intervalmean$steps, type='l',
	 xlab='5 minute interval', ylab='average number of steps')
```

## Imputing missing values

``` {R missing values}
sum(is.na(stepdata))
cleanstepdata <- stepdata[complete.cases(stepdata),]
dirtystepdata <- stepdata[!complete.cases(stepdata),]
cleanedstepdata <- merge(dirtystepdata[, -1], intervalmean, all.x=T)

allcleandata <- rbind(cleanstepdata,cleanedstepdata)

```

## Are there differences in activity patterns between weekdays and weekends?

```{R difference between weekdays and weekend}
allcleandata <- mutate(allcleandata, date = ymd(date), 
                       weekday = weekdays(ymd(date)),
                       day_type = ifelse(weekday == "Saturday" | weekday == "Sunday", "Weekend", "Weekday")
                       )
cleandata_weektype = aggregate(steps ~ interval+day_type, allcleandata,mean)

ggplot(data=cleandata_weektype, aes(interval, steps)) + geom_line() + 
  facet_wrap(~day_type, nrow=2) + ylab('Number of steps')
```

#### Some observations
+ during weekdays there seems to be a higher peak in number of steps in the early morning
+ during the day in weekends the number of steps on average is higher then during weekdays.
